### 2001-10-04 drafts/IST/PartB.tex

\documentclass[a4paper,11pt]{report}
\usepackage[latin1]{inputenc}
\usepackage{vmargin}

%% Marges
\setmarginsrb{2cm}{1cm}{2cm}{2cm}{1cm}{1cm}{1cm}{1cm}
%1 est la marge gauche
%2 est la marge en haut
%3 est la marge droite
%4 est la marge en bas
%5 fixe la hauteur de l'entête
%6 fixe la distance entre l'entête et le texte
%7 fixe la hauteur du pied de page
%8 fixe la distance entre le texte et le pied de page
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}
%\setcounter{tocdepth}{1}

%% XXX
\title{Proposal full title \\ Proposal Acronym}
\date{16/10/2001}
\begin{document}
\chapter*{Notice}
\begin{verbatim}
IST2001 - IV.3.3 Free software development: towards critical mass 
Objectives:

1.To foster in Europe a critical mass of development of free software released
  under GPL-compatible licenses. 
2.To make available European based support services for free software projects.

Focus:

1.Support services for free software developers at all stages of the project 
  lifecycle: project hosting certification, release, repositories, dissemination. 
2.Trials of free software development in the following fields: media technology 
  for personal users,co-operative information production and sharing, and usability. 
3.Socio-economic studies on technology assessment, economic assessment and derived 
  business models

Types of actions addressed: Trials and studies Accompanying Measures. 
Links with WP 2000: Further focused Action Lines IST 2000 - IV.3.4 and IV.3.5

more info : http://www.cordis.lu/ist/bwp_en5.htm
\end{verbatim}
\maketitle
\tableofcontents
\chapter*{some text}
%%%%%%%%%%%%%%%%%%%
\subsubsection{Context}
Most of free or open source software is collaboratively developped using an
hosting plateform. They are mostly web sites. The frontend integrates different
functionnalities that are available to project developpers and administrators
and linked them with each other. The backend is mainly implemented with
scripts. \\
Current hosting plaforms are implemented as a  web site that provides
functionnalities to developpers. The most common ones include  users and groups
of users management, project repository (usualy based on Concurent Versionning
System), per project web pages and document sharing, per project mailing lists
and web forums, bugtracking system, project management system (audit, task
management), global site management and information feeding. There is currently
four main web hosting facilities : SourceForge (http://sourceforge.net/),
Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
Savannah and SourceForge are the leaders.
\subsubsection{Mission}
Although hosting facilities are widely used in the freesoftware community the
existing solutions have some drawbacks, that refrain their wide adoption as
well in the Libre software community than in the software industry. \\
Due to a quick design, they are hard to distribute. They mainly are a front end
to a group of dedicated machines that provides a precise set of
functionnalities. Even if the acces is remote, all data is centralized as one
site physically hosts an huge number of projects. This is bad practice in terms
of data safety, data access and scalability. \\
Also due to the number of services and their inter-relations, it is not simple
to set up a new site that provides hosting. This is a real brake to a large
number of sites using this technology.  There is no complete solutions to
install easily any of the technologies described above. \\
They rely too much on managing procedures and human control. For example, a
site needs to be managed by people that acknowledge project creation. \\
At last the most active effort to build such a sytem has recently become
proprietary.  Therefore the produced work is not available. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{B3. Objectives}
\textit{This section, which should not exceed two pages, describes the
scientific/technological objectives of the proposal. They should be achievable
within the project, not through subsequent development, and should be stated in
a measurable and verifiable form. The Progress of the Take-Up action will be
measured against theses objectives in later reviews and assessments.\\
This section should clearly specify in addition to the technical objectives -
also the quantified business objectives and how they are measured}
\\ \\
to be written
\chapter*{B4. Contribution to program/Key action objectives}
\textit{This section, which should not exceed more than one page, describes how
the proposed Take-Up Action will contribute to the objectives of the program
and Key action. This can be done by describing how the proposal meets the
objectives and focus of the Action line which it addresses}
\\ \\
to be written
\chapter*{B5. Innovation}
\textit{This section, not exceeding two pages, describes the innovative
aspects of the proposed project. \\
Progress beyond the state of current practices within a user organization
should be described. The individual action steps must be described in the
context of the real business case. A separate description of this underlying
baseline project/target application - in terms of technical and business
parameters - must be included. Choice and justification for each of the new
practices including the supporting technologies (with comparison of not chosen
options) should be given (e.g. new methodologies, technologies, hardware and/or
software tools, subcontracted experts etc. -- whatever is applicable for the
experiment.}
\\ \\
to be written.
\chapter*{B6. Project workplan}
\textit{This section concisely describes the work planned to achieve the
objectives of the proposed Take-Up Action. The recommended length, excluding
the forms specified below, is up to 10 pages. An introduction should explain
the structure of the workplan and how the workplan must be broken into
workpackages (WPs) which should follow the logical phases of the projects'
life, and include management of the project and assessment of progress and
results. \\ \\
There must be a separate WP each for 'measurement and evaluation of results}
\\ \\ not yet written
\end{document}