### 2001-10-08 drafts/IST/PartA.tex
new objectives, and description of the work
svn path=/trunk/; revision=1483

\documentclass[a4paper,11pt]{report}
\usepackage[latin1]{inputenc}
\begin{document}
\section*{Notice}
This file will be generated by Protool. Thus layout does not matter. It will
juste be cut\&paste.
\section*{A1 Administrative Overview}
\subsection*{Summary}
\begin{tabular}{|c|c|}
\hline
Proposal Acronym &  \\ \hline
Proposal Number &  \\ \hline
Thematic Priorities & IST-2001-4.3.3. \\ \hline
Call identifier & \\ \hline
Proposal Date & 16 10 2001 \\ \hline
Proposal full name & \\ \hline
\end{tabular}
\subsection*{Abstract (1000 characters)}
\textit{The proposal abstract should be a very short and precise presentation
of the main features of the proposal. Why is it proposed and what problem is it
solving? What are the objectives? How will the objectives be achieved? What
results are expected? This proposal abstract will be used together with the
proposal summary description in form A2 in the evaluation process and in
communications about the proposals to the interested parties (Commission
services and programme committees, etc.). Please use plain typed text, avoiding
formulae and other special characters. If the proposal is written in a language
other than English, please include in form A1 an English version of the
abstract.} \\ \\ \\
In almost two years, the biggest barriers for Free Software development have
been raised by hosting facilities which provide most of the mandatory tools
that free software developers need for everyday tasks such as Bug tracking,
mailing lists, Project management... But although there are
many hosting facilities around the world, no software package is designed to
setup and maintain a set of machines dedicated to this purpose. The NAME
project will provide a generic solution that can be used and adapted by all of
them.
NAME will of course provide the same services as current facilities but
will focus an the ease of configuration and installation that all current ones
lack.
\\ \\ 
680 characters...
\begin{tabular}{|c|c|}
\hline
Duration in months & \\ \hline
Keyword & BESTP \\ \hline
\end{tabular}
\section*{A2 Proposal Summary}
\textit{This form, should be filled in by the co-ordinator only, for single
participant proposals by the principal contractor. You should not use more than
3,500 characters. The proposal summary should, at a glance, provide the reader
with a clear understanding of the proposal objectives and how the objectives
will be achieved, and their relevance in the context of the objectives of the
specific programme. This summary may be used as an alternative to the proposal
abstract, as the description of the proposal in the evaluation process and in
communications to the programme management committees and other interested
parties. It must therefore be short and precise. Please use plain typed text,
avoiding formulae and other special characters. If the proposal is written in a
language other than English, please include in form A2 an English version of
the proposal summary.} \\
\subsection*{Objectives (max 1000 characters)}
The NAME project aims at developping a complete hosting facility. It composed
of two subprojects. \\
The first one is the NAME project that will be a next generation hosting
facility. It will provide a web frontend to a number of tools are now mandatory
for software development (user/groups rights management, bug tracking, mailing
lists and forums...). The main idea between its design is to make it
transparently distributed. At each step of its development, we will focus on
the ease of installation and configuration.
The second one is the "CoopX" project that will provide a generic description of
an hosting facility. With that tool we want to allow different kind of facility
to communicate each other. It will help to develop tools to facilitate
migrations between hosting facilities.
\\ 800 chars need some check and some other stuff:w
\subsection*{Description of the work (max 2000 characters)} to be written
%SAVANNAH
The hosting facility is composed of two main parts, the frontend and the
backend. The first one is the code base that interacts with the user and that
will implement services that are not provided by external software. The backend
will link the Database (filled by the frontend) to external software which
provide services such as DNS, mailing lists or CVS. Concurrently the packaging
of the solution will allow the validation and testing of each milestone.
% 450
%COOPX
\subsection*{Milestones and expected result (max 500 characters)}
%%% HELP HELP
\end{document}